### 2010-04-26 conference-paper/proposal.tex
First draft of conference submitted

\documentclass{conference}

% For graphics
\usepackage{graphicx}

% Bibliography
\bibliography{papers,websites,patents}

% Begin document
\begin{document}
{\center \Large \bfseries \MakeUppercase{Using traffic analysis to identify Tor} \par}

{ \center
  \begin{tabular*}{\textwidth}{cccc}
  John Barker & Peter Hannay & Christopher Bolan & Patryk Szewczyk \\
  jebarker@our.ecu.edu.au & p.hannay@ecu.edu.a & c.bolan@ecu.edu.au & p.szewczyk@ecu.edu.au \\
  \end{tabular*}
}

\begin{abstract}

Traditional attacks against anonymous systems aim to uncover the identities of those involved. However, a more likely goal of attackers is to block or degrade the network itself, discouraging participation and forcing vulnerable users to communicate using less secure means. Since anonymous networks operate on known protocols and employ strong encryption it is difficult to distinguish them from regular traffic. This proposal describes a method for identifying traffic belonging to anonymous networks by examining their communication patterns.

\end{abstract}

\section{Introduction}

The Internet has revolutionised the political sphere in providing a platform for the publication of speech to a far greater audience than was ever available before digital communications systems \parencite{Bonchek:1997p3455}. The publication of less desirable political speech, criticism or challenging ideas still carries with it great risk, with numerous publications leading to arrest. This has spawned headlines such as \emph{Egypt arrests another blog critic} \citeyear{website:egypt-arrests}, \emph{Chinese couple sue Yahoo! in US over torture case} \parencite{website:china-yahoo-torture}, \emph{Vietnam bloggers arrested over China shirt protest} \parencite{website:vietnam-bloggers-arrested}, \emph{23 Journalists/Bloggers Arrested in Iran, Including Head of Top Group} \citeyear{website:iran-bloggers-arrested} and \emph{Blogger Arrests} \parencite{website:blogger-arrests}. The Internet, like conventional media is still vulnerable to censorship by oppressive governments and malicious attackers \parencite{Crandall:2007p6165,Karlin:2009p6166}.

In response to these threats, a number of systems have been proposed which use cryptography to provide censorship resistance and anonymity, including The Second Generation Onion Router (Tor). Tor is an anonymous network which uses Onion Routing to provide low latency anonymity to network participants \parencite{Dingledine:2004p314}. It also aims to provide resistance against censorship based blocking attempts \parencite{Dingledine:2008p1542}.

Although Tor provides censorship resistance and a certain amount of privacy, it may be possible to use traffic analysis techniques to automatically identify Tor connections and target them for attacks. Since Tor relies on large numbers of users for the protections it provides, attacks that affect the reliability of the network can discourage participation, rendering it`s protections useless. This paper proposes a statistical methodology for identifying Tor traffic which may be used for such attacks in the future.

\section{Significance}

Tor`s availability, unrestrictive license and platform portability has made it a popular anonymous network with around one thousand eight hundred users at the time of writing \parencite{website:tor-anonymity-online}. It has been the subject of many technical papers proposing traffic analysis attacks, however the primary objective has historically been attacks that reveal the identity of participants. As stated in \textcite{Murdoch:2005p325} ``The principal objective of an adversary attacking an anonymous communication system is to link the initiators of connections with their respective communication partners and vice versa''.

The traffic analysis techniques proposed will attempt to classify network streams belonging to Tor users to distinguish them from regular encrypted communications.

\section{Research Questions}

The ability of Tor to connect on any port and it`s usage of strong encryption mean that recognizing traffic belonging to Tor is not straightforward. However it is possible that the communications patterns exhibited by Tor nodes are recognizeable using statistical methods. This raises the question: is it possible to distinguish the traffic created by these networks, from encrypted traffic produced by conventional software? This question can be broken down further:

\begin{enumerate*}
\item Can the traffic be classified using automated matching techniques?
\item Do these networks have characteristics that make them readily distinguishable using heuristics based matching?
\item Do they have characteristics that make them distinguishable using a machine learning technique?
\item How long does a user have to be connected to the network before a confident match can be made?
\end{enumerate*}

\section{Background Information}

The first anonymous network system was proposed in \textcite{Chaum:1981p296}. This proposal included the use of public key cryptography and a centrally located server known as a `mix'. By wrapping the information and address of a destination in a message and encrypting it with the mix`s public key, only the mix system is able to read the message.

Onion routing was first described in a patent by \cite{Michael:2001}. Like the mix network it uses public key cryptography to provide anonymity, it differs in that messages are repeatedly encrypted then passed through several nodes in a network known as onion routers. Each of these routers is capable of removing a single layer of encryption to reveal delivery instructions. This layering prevents individual nodes obtaining the identity of the sender, receiver and the contents of the message.

\subsection{The Second Generation Onion Router (Tor)}

The Second Generation Onion Router (Tor) uses a form of Onion Routing which employs telescoping circuits to provide low latency anonymity \parencite{Dingledine:2004p314}. A connection through the Tor network constitutes a \emph{circuit}, which hops over several Tor nodes known as \emph{relays}. Relays that allow connections directly from Tor clients are known as \emph{bridges} and relays that deliver messages from the frontier of the Tor network are known as \emph{exit nodes}. Figure \ref{tor-network} shows how a client connects to the Tor network.

\begin{figure}[H]
\center \includegraphics[scale=0.8]{tor-network-diagram}
\caption{The Tor Network}
\label{tor-network}
\end{figure}

When a relay joins the Tor network, it trades symmetric cipher keys with its neighbours using public key cryptography. These keys are used to encrypt communications and are destroyed when the session is closed to prevent replay attacks.

After it was noticed that Tor was being used to circumvent censorship as well as for anonymity, a proposal was developed to improve its blocking resistance \parencite{Dingledine:2008p1542}. Many of the suggestions in this proposal have been implemented, but many still remain.

\section{Related Work}

When considering the use of traffic analysis for Internet communications, three primary techniques are used. These are: exact matching, heuristic matching and machine learning \parencite{Zhang:2009p1188}. Exact matching techniques are ineffective in recognizing Tor communications as Tor can operate on any available port and uses encryption to hide communications. This leaves heuristic and machine learning matching techniques.

\subsection{Heuristic Techniques}

The increasing burden of peer to peer (P2P) applications on campus networks, and their shift to encryption motivated the development of heuristics based techniques for identifying P2P traffic. Early techniques include identifying known properties of P2P networks such as often communicating using both the UDP and TCP protocols simultaneously and using a solitary connection to transfer high volumes \parencite{Karagiannis:2004p6400}. Similar techniques have been developed in \parencite{Perenyi:2006p6325} and \parencite{John:2008p1376}, both of which attempt to improve matching accuracy by expanding the scope of matching parameters and eliminating false positives using exact matching techniques.

Heuristics based matching techniques have also been used to identify traffic belonging to network worms \parencite{Lazarevic:2003p6450} and the Skype protocol \parencite{Ehlert:2006p2396}.

\subsection{Machine Learning Techniques}

There are a vast number of traffic analysis papers that utilize machine learning techniques for the purpose of traffic classification, demonstrating a wide variety of algorithms for this purpose. The first use of machine learning to categorise traffic flows appears in \textcite{McGregor:2004p3826}. A detailed analysis of the attributes that can be used for machine learning and an attempt at coarse grained classification using an Expectation-Maximisation (EM) algorithm are demonstrated. The same technique is also employed in \textcite{Soule:2004p3817} using histograms for finer grained classification. The EM algorithm is again used in \textcite{Zander:2005p6212} and \textcite{Erman:2006p3825}, with the latter also comparing this algorithm favourably against a Na\"ive Bayes classifier.

\textcite{Moore:2005p3827} demonstrates the usage of a supervised Na\"ive Bayes algorithm to classify traffic flows. This paper focuses on many of the most commonly used Internet protocols while \textcite{Bonfiglio:2007p6453} uses the technique for identifying traffic belonging to the commercial Skype application.

\textcite{Herrmann:2009p1189} uses Bayesian networks to fingerprint visited websites accessed through Privacy Enhancing Technologies (PET), including Tor. This technique performed poorly when applied to the Tor network, but it suggests that Tor traffic has particular characteristics that distinguish it from many existing PETs. It makes a particularly useful observation: ``The most frequent packet sizes in the Tor traffic dumps are, in descending order, 1500, −52, −638, 52, 638 and 1150 bytes, accounting for 87.6\% of all Tor packets.''

Hidden Markov Models (HMMs) are first used as a traffic analysis technique in \emph{HMM profiles for network traffic classification} \parencite{Wright:2004p3860}. The primary identification characteristics for use with this algorithm are packet size and inter-packet arrival times. With refinement this algorithm is used with increasing accuracy in \textcite{Wright:2006p322} and \textcite{Dainotti:2008p1435}. HMMs are also used in \textcite{Bernaille:2005p6205} to discover distinguishing characteristics of traffic flows, rather than specifically as a classifier.

Clustering algorithms group observations into subsets based on similar characteristics. They have been used in a number of traffic classification techniques with the K-Means clustering technique being the most prominent. K-Means cluster analysis appears in Bernaille et al. (2006), \textcite{Erman:2007p3764}, \textcite{Erman:2007p6206}. The use of the Density Based Spatial Clustering of Applications with Noise (DBSCAN) algorithm appears in \textcite{Erman:2006p3766}, alongside K-Means and Autoscan algorithms.

Other algorithms used for traffic classification include Nearest Neighbour and  Linear Discriminant Analysis \parencite{Roughan:2004p3823}, Normalized Threshold \parencite{Crotti:2007p3824}. The use of the Gaussian Mixture Model to identify applications and identities inside SSH tunnels was demonstrated in \textcite{Dusi:2008p6254}.

\subsection{Comparing Techniques}

It is difficult to say which technique would be most suitable for identifying Tor traffic. In the literature, no particular consensus has been reached and no two papers use the same baseline data set for their technique and so no comparison can be made \parencite{Kim:2007p3867}. Although there have been some comparisons of machine learning algorithms \parencite{Williams:2006p3849,Mohd:2009p6484}, no paper has published their data set due to privacy concerns. It is clear that choosing the right technique will require some experimentation and comparison of the effectiveness of each algorithm.

\section{Theoretical Framework}

The proposed experiment will employ a quasi-experimental research method, employing techniques common to traffic analysis papers. This includes the capturing of packet traces, development of a matching technique/algorithm and documentation of efficacy and results. The first phase of the experiment covers the data capturing phase, while the second phase concludes the traffic analysis technique.

\subsection{Assumptions}

It is assumed that the usage patterns exhibited by individual users will be smaller than the communications characteristics that will lead to the identification of anonymous and censorship resistant networks. Thus there is no need to obtain a large sample of regular network traffic from varying user profiles.

As of the current implementation, Tor network traffic is readily distinguishable by looking at the handshake packets. It is likely that this weakness will be addressed in a future version of the Tor protocol as it is recognised as a design goal in \textcite{Dingledine:2008p1542}. For this reason, this proposal focuses on traffic analysis techniques that are content agnostic.

\subsection{Variables}

The data capturing stage will be affected by a number of variables that will influence the accuracy of the chosen matching algorithms. These include:

\begin{enumerate*}
\item System Performance
\item Network Performance
\item Application Protocol
\item Quality of the Anonymous Network
\item Caching
\end{enumerate*}

An attempt will be made to reduce the impact of these variables by capturing packets generated by an isolated simulation of the Tor network, ensuring only vital network applications are running and executing software inside a virtual machine which can be rolled back to a known snapshot \parencite{website:snapshot-definition}.

\section{Materials and Methods}

The experiment will consist of two phases, first a simulation phase which involves the capture of packets belonging to a Tor network, the second will be a comparison and documentation of the results of using several techniques to identify Tor traffic.

Figure \ref{network-diagram} shows the placement of applications and relevant data flows for the simulation phase.

\begin{figure}[H]
\center \includegraphics[scale=0.7]{network-diagram}
\caption{Network Diagram}
\label{network-diagram}
\end{figure}

\subsection{Methodology}

\subsubsection{Data Collection}

The data collection phase will be conducted by running a series of simulations designed to mimic real world traffic. These simulations will be controlled so that the data captured will only contain Tor traffic in the training and testing data sets, and non Tor encrypted traffic in the data set for determining the false positive rate.

\subsubsection{Algorithm Selection}

Once the data collection phase is complete, a number of different algorithms will be chosen to classify Tor traffic. These will be compared using a number of criteria to rate effectiveness. These will include the accuracy of the matching technique and the number of packets needed to make a confident identification.

\printbibliography[title=REFERENCES]

\end{document}

% vim: fdm=syntax