### 2010-10-24 special-topic/paper.tex
Beginning of special topic paper

\documentclass{ecuthesis}

% For graphics
\usepackage{graphicx}

% Bibliography
\bibliography{fullrefs,websites,patents,papers}

\begin{document}
{\center \Large \bfseries \MakeUppercase{Using traffic analysis to identify Tor} \par}

{ \center
  \begin{tabular*}{\textwidth}{cccc}
  John Barker & Peter Hannay & Christopher Bolan & Patryk Szewczyk \\
  jebarker@our.ecu.edu.au & p.hannay@ecu.edu.a & c.bolan@ecu.edu.au & p.szewczyk@ecu.edu.au \\
  \end{tabular*}
}

\input ../thesis/experiment

\printbibliography[title=REFERENCES]

\end{document}

% vim: fdm=syntax wrapmargin=80