### 2010-10-24 special-topic/paper.tex
Work on special topic paper, experiment notes, minted syntax highlighting, fix to thesis class

\documentclass{ecuthesis}

% For graphics
\usepackage{graphicx}

% Bibliography
\bibliography{fullrefs,websites,patents,papers}

% Setup the variables for the title page
\title{Using traffic analysis to identify The Second Generation Onion Router}
\author{John Barker}
\department{School of Computer and Security Science}
\school{Edith Cowan University}
\degree{Computer Science (Honours)}
\email{jebarker@our.ecu.edu.au}
\studentid{0991300}
\supervisor{Dr Andrew Woodward}
\supervisor{Mr Peter Hannay}
\supervisor{Mr Patryk Szewczyk}

\begin{document}
\maketitle

\chapter{Introduction}

\section{Background}

This document describes the procedures for configuring a small network of
computer systems to emulate a small sample of web based traffic using both the
HTTPS protocol and the Tor protocol. All traffic between the systems is captured
for the purpose of traffic analysis.

\input ../thesis/experiment

\chapter{Appendix}

\section{Manager Script}
\inputminted{ruby}{../experiment/simulation/manager/run_simulation.rb}

\section{Client Script}
\inputminted{ruby}{../experiment/simulation/client/client.rb}

\section{Selenium-RC Script}
\inputminted{bash}{../experiment/simulation/client/selenium-rc-with_tor.sh}

\section{Selenium-RC Script (with Tor)}
\inputminted{bash}{../experiment/simulation/client/selenium-rc.sh}

\section{Hosts file}
\inputminted{bash}{../experiment/hosts}

\section{Apache Site Configuration}
\inputminted{apacheconf}{../experiment/simulation/site/000-default-ssl}

\section{Tor Setup Script}
\inputminted{bash}{../experiment/simulation/tor/make_private_network.sh}

\section{Tor Running Script}
\inputminted{bash}{../experiment/simulation/tor/run_private_network.sh}

\printbibliography[title=REFERENCES]

\end{document}

% vim: fdm=syntax