### 2011-02-26 thesis/experiment.tex
Rewriting of some bits of the equipment section

\chapter{Materials and Methods}

\section{Instruments and Apparatus}

The experiment described requires a number of computer systems and associated
hardware to run the simulation and capture appropriate data. This equipment is
described in the following chapter.

\begin{enumerate}
  \item Hub/Managed Router

  Most modern network appliances used for routing, even those labelled as hubs,
  include switching capabilities, which means that packets routed are only
  delivered to the addressed host. This makes eavesdropping difficult. For
  the purpose of this experiment a simple hub or repeater is required. A hub
  does not examine packet delivery information, it simple broadcasts all packets
  received to every connected port, allowing for passive eavesdropping without
  complicated network configuration \parencite{website:hub-reference}.

  \item 'tcpdump' Software

  The tcpdump software captures network packets that are transferred using
  network interfaces attached to the current hardware \parencite{:2009cr}.  It is
  available on most operating systems and supports a flexible filtering language
  for capturing only the specific packets desired. The format of output produced
  by tcpdump is well documented \parencite{:nx} which makes it ideal for
  transformation and processing by third party tools.

  \item Tor Software

  The Tor software is available on a number of operating systems and the Tor
  project makes packages available for the Ubuntu operating system.

  \item Weka Software

  Weka is data mining software which includes tools for classification of data
  sources using a wide variety of machine learning algorithms
  \parencite{Hall:2009p7662}. It supports processing of an input
  Attribute-Relation File Format (ARFF), a file which contains ASCII text
  describing a list of items with common attributes. Items can be given an
  attribute which provides a nominal classification for verification purposes.

  \item VirtualBox Software

  VirtualBox software allows multiple guest operating systems to run in an
  isolated virtual environment on a host operating system \parencite{:fk}.
  It is free software and supports both snapshots and a remote control facility
  which make it useful for running a controlled experiment \parencite{:uq}.

  \item Ruby Software

  Ruby is an object orientated, dynamic programming language which is useful for
  rapid prototyping and is supported by a large community that has developed
  powerful add on software in the form of Ruby Gems \parencite{:2010uq}. These
  Gems include support for the Selenium browser testing framework, process
  monitoring, dynamic web content generation and automated testing.

  \item Ubuntu Linux Software

  The Ubuntu Linux operating system is freely available Linux operating system
  available for desktop personal computers \parencite{:2010ly}. It comes
  packaged with the 'tcpdump' software and can be configured in a reliable and
  consistent way with typical set of POSIX tools and the 'apt-get' package
  manager \parencite{Pereira:2011uq}.

  \item Selenium Software

  Selenium is a web testing framework which allows automated testing of web pages
  through any Javascript enabled web browser \parencite{:2010ys}. A web browser
  can be controlled remotely using the Selenium Remote Control Server component,
  which in turn can be controlled with a number of programming languages including
  Ruby. Using Selenium the typical interactions a user makes when viewing a
  website can be simulated.

  \item Firefox Software

  Firefox is a web browser developed by the Mozilla organization and freely
  available for most desktop personal computers \parencite{Foundation:2011fk}.
  It's strong standards compliance \parencite{Hammond:2010fk} and popularity
  \parencite{:2010kx,:2010vn} on the simulation platform make it an ideal
  candidate for this experiment.

  \item OpenSSH Software

  OpenSSH is a freely available suite of connectivity tools
  \parencite{:2010zr,:ve} that supports remote command execution
  \parencite{Tucker:2010ly, :2010zr}. The client component of OpenSSH is
  installed by default on the Ubuntu operating system and the server component
  is easily installed using the apt-get package management system.

  \item Apache Software

  The Apache web server is the most prevalent web server used on the Internet
  \parencite{2010:dq}. The Apache web server can serve static web content over
  both the HTTP and HTTPS protocols \parencite{Foundation:2011uq}. Combined with
  software such as Sinatra it can serve dynamic web content.

  \item Sinatra Software

  Sinatra is a simple framework for creating dynamic web applications using the
  Ruby programming language \parencite{:2010bh}. Sinatra requires very little
  configuration and development time to develop dynamic web applications.

\end{enumerate}

\section{Procedure}

The experiment involves two stages, an initial configuration stage to prepare
equipment, install software, configure and test the design and the final stage
of executing the simulation to capture packet traces.

\subsection{Hardware Configuration}

Four identical desktop computer systems were prepared and arranged as shown
in figure \ref{physical-setup}.

\begin{figure}[H]
  \centering\includegraphics[scale=0.7]{physical-setup}
  \caption{Physical Setup}
  \label{physical-setup}
\end{figure}

\subsection{Software Configuration}

Table \ref{table:hosts} shows the role of each system in the experiment as well
as the edition of operating system in use and the IP address they were
configured with on the network.

\begin{tabular}{lrr}
  \toprule
  Role & Operating System Version & IP Address\\
  \midrule
  Client Host & Desktop & 192.168.0.100\\
  Server Host & Desktop & 192.168.0.001\\
  Tor Host & Desktop & 192.168.0.102\\
  \midrule
  Sniffing System & Server & 192.168.0.200\\
  \midrule
  Client Guest & Desktop & 192.168.0.16\\
  Server Guest & Server & 192.168.0.17\\
  Tor Network Guest & Server & 192.168.0.18\\
  \bottomrule
  \label{table:hosts}
\end{tabular}

Figure \ref{network-diagram} shows the placement of applications and relevant
data flows.

\begin{figure}[H]
  \centering\includegraphics[width=\linewidth]{network-diagram}
  \caption{Network Diagram}
  \label{network-diagram}
\end{figure}

\subsection{Standard OS Installation Procedure}
\label{section:os_install}

All systems in the experiment use a variation of the Ubuntu Linux operating
system. The installation process followed is detailed below:

\begin{enumerate*}
  \item Configure system BIOS to boot from CD-ROM
  \item Insert Ubuntu operating system installation CD
  \item Select the language you wish to use for installation (English) and click
    'Forward'
  \item Select the region (Australia/Perth) and click 'Forward'
  \item Select an appropriate keymap (USA) and click 'Forward'
  \item Select 'Erase and use the entire disk' and click 'Forward'
  \item Enter the user name (John Barker), select 'Log in automatically' and click 'Forward'
\end{enumerate*}

At this point the installation will begin and should take about 35 minutes.
'Log in automatically' is selected to ensure that the experiment will continue
even if power is lost. Once installation is complete remove the CD and reboot
the system. Once booted the system is ready for further configuration.

Each system is given a statically assigned IP address, this is done by editing
the file \verb+/etc/network/interfaces+ and changing the line
\verb+iface eth0 inet dhcp+ to:

\begin{lstlisting}[language=sh]
iface eth0 inet static
  address 192.168.0.16
  netmask 255.255.255.0
  network 192.168.0.0
  broadcast 192.168.0.255
\end{lstlisting}

Substituting \verb+eth0+ with the physical network interface identifier as
appropriate, and the address \verb+192.168.0.16+ with the desired value.

\subsection{Sniffing System}

The sniffing system is prepared as described in \ref{section:os_install} with an
Ubuntu 10.04 server CD. The base install of Ubuntu server has all the required
packages included so no additional software installation is required.

An init.d script was created to execute the tcpdump program with a specific
filter string after the system had powered on. All capture files were limited to
1mb in size and put into a time stamped directory. To ensure that only relevant
packets were collected the filter string was tuned for each phase of the
experiment. The filter string for each phase is shown below.

Phase 1:
\begin{lstlisting}[language=sh]
tcp port 443
\end{lstlisting}

Phase 2 and 3:
\begin{lstlisting}[language=sh]
tcp portrange 5000-5015
\end{lstlisting}

The capture script appears below with the filter string as required for phase 2
and 3:

\begin{lstlisting}[language=sh]
#!/bin/bash
export capture_date=`date`
cd /home/john/ssl-captures
mkdir "./${capture_date}"
tcpdump -vv -C 1 -i eth0 -w "./${capture_date}/capture.dat" tcp portrange 5001-5015
\end{lstlisting}

\subsection{Host Systems}

The host systems all run an identically configured Ubuntu 10.04 desktop
operating system, except for the IP address which is configured as detailed in
\ref{table:hosts}. To support the guest operating systems, VirtualBox is
required. The procedure for installing VirtualBox is detailed below:

\begin{enumerate*}
  \item Edit the file \verb+/etc/apt/sources.lst+ and add the line:
\begin{lstlisting}[language=sh]
deb http://download.virtualbox.org/virtualbox/debian lucid non-free
\end{lstlisting}
  \item Install the GPG key for the Virtual Box packages:
\begin{lstlisting}[language=sh]
wget -q http://download.virtualbox.org/virtualbox/debian/oracle_vbox.asc -O- | sudo apt-key add -
\end{lstlisting}
  \item Then install VirtualBox 3.2 with the command:
\begin{lstlisting}[language=sh]
sudo apt-get update && sudo apt-get install virtualbox-3.2
\end{lstlisting}
\end{enumerate*}

The guest operating system images can then be created or imported from an
external disk.

\subsubsection{Server Host}

One of the host systems is set aside as a management system, which is
responsible for ensuring the experiment runs and restarts it if there are any
problems. On this machine a private SSH key is created, and copied to the
authorized keys file on all other hosts. This allows the management system to
connect to all the other systems and issue commands without user interaction.

The management script runs on the management system to watch the experiment's
progress. This is a simple Ruby script which uses SSH to control the other Host
systems and opens a socket which waits for a connection that indicates that the
experiment has finished. Every time it starts the experiment, it rolls back
each virtual machine to a preconfigured snapshot. If the experiment takes too
long, say because a virtual machine hangs or something is wrong, it will be
started without waiting for the finish condition. Figure
\ref{manager-flow-diagram} contains a diagram which outlines the basic flow of
the manager script.

\begin{figure}[H]
  \centering\includegraphics[scale=0.7]{manager-flow}
  \caption{Manager Script Flow Chart}
  \label{manager-flow-diagram}
\end{figure}

Note that restarting each snapshot ensures that the system clock of each guest
operating system is synchronized with that of the host operating system. The
Tor network is sensitive to differences in time between hosts and this ensures
correct operation.

\subsection{Guest Systems}

\subsubsection{Client System}

The client system runs Ubuntu 10.04 Desktop Operating System as well as a
number of applications that allow it to perform automated web browsing. Tor is
configured on this system as well for the Tor part of the experiment but is
initially disabled.

The software installed on this system includes:

\begin{itemize*}
  \item Ruby
  \item Rubygems:
    \begin{itemize*}
      \item Selenium
      \item DaemonController
      \item RSpec
    \end{itemize*}
  \item Selenium-RC
  \item Firefox web browser
  \item Tor
\end{itemize*}

The client operation is controlled by a Ruby script which ensures that the
Selenium-RC server is running, executes a number of Selenium browser
simulations using RSpec and reports it's success by connecting to an open
socket on the management system.

There are 170 simulations which emulate simple website interactions which have
different traffic profiles against 30 different websites on the Server system.

For the non-SSL portion of the experiment, Selenium is configured to use a
special proxy which automatically accepts self signed certificates. Typical
interactions with SSL encrypted websites are signed by a certificate authority,
so no user interaction is required to begin browsing.

\paragraph{Selenium}

SSL + Tor over selenium.

Generate a selenium firefox profile. First launch selenium in interactive mode
specifying a proxy:

\begin{lstlisting}[language=sh]
java -jar $selenium.jar -interactive
\end{lstlisting}

This will start a prompt to which you can edit commands, you must launch a
browser to get selenium to create a firefox profile directory.

\begin{lstlisting}[language=sh]
cmd=getNewBrowserSession&1=*firefoxproxy&2=http://www.google.com
\end{lstlisting}

Generate a firefox profile, then edit the proxy.pac file to

\begin{lstlisting}[language=sh]
function FindProxyForURL(url, host) {
   if(shExpMatch(url, '*/selenium-server/*')) {
       return 'PROXY localhost:4444; PROXY proxy.ourdomain.dom:8080';
   } else if (shExpMatch(host, '*dev.ourdomain.dom*')) {
       return 'DIRECT';
   } else if (shExpMatch(host, '*qa.ourdomain.dom*')) {
       return 'DIRECT';
   } else if (shExpMatch(host, '*staging.ourdomain.dom*')) {
       return 'DIRECT';
   } else {
       return 'PROXY proxy.ourdomain.dom:8080';
   }
}
\end{lstlisting}

or

\begin{lstlisting}[language=sh]
function FindProxyForURL(url, host) {
   if(shExpMatch(host, '*torsim')) {
       return 'PROXY localhost:8118';
   }
   return 'PROXY localhost:4444; DIRECT';
}
\end{lstlisting}

This makes sure that your proxy (Tor) gets used for outgoing connections, and
the selenium proxy is used for running tests.

To get SSL certificates accepted, you must manually accept them permanently for
each host. This will update the \verb+cert_override.txt+ and \verb+cert8.db+
files.

Tor must have the following setting supplied:

\begin{lstlisting}[language=sh]
ExitPolicyRejectPrivate 0
\end{lstlisting}

Since we are using a 192.* network. - Not true, testing mode flag sets this.

But the tor relays must be able to resolve names using DNS, if you use socks4a protocol.

\subsubsection{Server System}

The host system runs an apache web server which serves two sets of sample
websites, static and dynamic. The static websites are served each from a
different numbered directory and accessible over a different virtual host. This
means that when a web browser is pointed at the directory
\verb+site1.static.torsim+ the files in
\verb+/var/www/thesis/experiment/simulation/static/site1+ are served to the
browser.

For the SSL part of the experiment, only the HTTPS port 443 is accessed and all
information is sent to the browser using encryption.

The dynamic sample websites change subtly with each attempted access. There are
10 sample websites each a slightly different Ruby Sinatra application.

\subsubsection{Tor Simulation}

The Tor Simulation Guest is responsible for emulating a Tor network. It runs
three Tor directory authorities as well as fifteen relays.

Each Tor instance on the Tor Simulation system is given it's own separate
directory for configuration, logging and temporary files.

Deploy the experiment files.

Install Tor:

\begin{enumerate*}
  \item Edit the file \verb+/etc/apt/sources.lst+ and add the line:
    \begin{lstlisting}[language=sh]
deb http://deb.torproject.org/torproject.org lucid main
    \end{lstlisting}
  \item Install the GPG key for the Tor packages:
    \begin{lstlisting}[language=sh]
gpg --keyserver keys.gnupg.net --recv 886DDD89
gpg --export A3C4F0F979CAA22CDBA8F512EE8CBC9E886DDD89 | sudo apt-key add -
    \end{lstlisting}
  \item Install Tor:
    \begin{lstlisting}[language=sh]
sudo apt-get update && sudo apt-get install tor
    \end{lstlisting}
\end{enumerate*}

\subsubsection{Client System}

Install Ubuntu 10.04 Desktop as in the Host system directions, leaving out the
Virtual Box installation.

\begin{enumerate*}
  \item Deploy the experiment files onto the client system.
  \item Install Ruby
  \item Install the bundler ruby gem with the command:
    \begin{lstlisting}[language=sh]
gem install bundler
    \end{lstlisting}
  \item Install the required client ruby gems by executing the bundler from the client directory in the deployed experiment files.
    \begin{lstlisting}[language=sh]
bundle install
    \end{lstlisting}
  \item Ensure that the client script is executed at system startup.
\end{enumerate*}

\subsubsection{Configuring the Management Script}

Once all systems are deployed, the management script needs to be altered to

\subsection{Execution}

Running the experiment is a matter of ensuring all Host systems are powered on,
if everything is configured correctly the experiment will continue to run until
stopped manually.

\subsection{Processing of Results}

The experiment was run over the course of 8 weeks yielding three sets of sample
data.

\begin{tabular}{lrr}
  \toprule
  Phase & Size (bytes) & Number of sessions\\
  \midrule
  HTTPS & 2326328 & 236659\\
  HTTP over Tor & 2869624 & 168876\\
  HTTPS over Tor & 1591112 & 95203\\
  \bottomrule
  \label{table:datasets}
\end{tabular}

The data captured was in the form of 1mb capture files which were recombined
with mergecap and processed by NetAI to produce ARFF format files for use by
Weka. Each data set were given the nominal classification 'http',
'http\_over\_tor' and 'https\_over\_tor', all other nominal attributes were
removed.

% vim: fdm=syntax tw=80