### 2011-04-17 thesis/results.tex
Revised questions a little, added paragraph to limitations

\chapter{Results}

\section{Results}

The classification results are tabulated below. The best performing algorithm
random forest was able to classify HTTP over Tor with 93.7\% accuracy and a
false positive rate of 3.7\%. HTTPS over Tor was more easily identified with a
97.7\% accuracy and low false positive rate of 0.3\%.

The table includes the true positive and false positive rate rate of each algorithm and
include the area under the receiver operating characteristic (ROC) curve. The
ROC is a tool used for evaluating the usefulness and comparing different
matching algorithms and is often used to compare Internet traffic
classification techniques \parencite{citeulike:167557,Nguyen:2008p3837}.

\begin{table}[H]
  \begin{tabular}{lrrr}
    \toprule
    Algorithm & True Positive Rate & False Positive Rate & ROC Area \\
    \midrule
    Random Forest\\
    \midrule
    HTTPS & 0.957 & 0.036 & 0.99\\
    HTTP over Tor & 0.937 & 0.037 & 0.986\\
    HTTPS over Tor & 0.977 & 0.003 & 0.999\\
    Weighted Avg. & 0.954 & 0.03 & 0.99\\
    \midrule
    j4.8 With 10 fold cross validation\\
    \midrule
    HTTPS & 0.951 & 0.04 & 0.989\\
    HTTP over Tor & 0.978 & 0.043 & 0.98\\
    HTTPS over Tor & 0.97 & 0.007 & 0.992\\
    Weighted Avg. & 0.964 & 0.018 & 0.986\\
    \midrule
    Random Tree\\
    \midrule
    HTTPS & 0.961 & 0.046 & 0.963\\
    HTTP over Tor & 0.906 & 0.04 & 0.94\\
    HTTPS over Tor & 0.955 & 0.01 & 0.972\\
    Weighted Avg. & 0.941 & 0.037 & 0.957\\
    \midrule
    Adaboost\\
    \midrule
    HTTPS & 0.95 & 0.001 & 0.975\\
    HTTP over Tor & 0.999 & 0.324 & 0.838\\
    HTTPS over Tor & 0 & 0 & 0.777\\
    Weighted Avg. & 0.785 & 0.109 & 0.891\\
    \bottomrule
  \end{tabular}
  \caption{Results}
  \label{table:results}
\end{table}

\chapter{Discussion of Results}

\section{Limitations}

The classifiers developed within this experiment have been trained against
a relatively small sample set of simulated data, which may only represent
a small fraction of the width and breadth of conditions available on a wider
area network. The characteristics that allow for classification of Tor network
may be masked by the natural variability introduced by a heterogeneous network.

The real Tor network consists of over two hundred thousand computer systems,
with over two thousand routers \parencite{TorStatus:2011fk}, all with different
operating systems, different hardware, different configurations and available
bandwidth. All the variables that were controlled for during this experiment
are present in the real network, and may be amplified by it's greater size.
Without further research it is difficult to say if the technique demonstrated
here will apply in the wild.

\section{Addressing Research Questions}

\begin{enumerate*}
  \item{Can Tor traffic be distinguished from other network traffic that uses
    similar or the same encryption and protocols?}

  Yes, the classification algorithms trialled were all able to identify Tor traffic
  with reasonable accuracy.

  \item{Can Tor traffic be identified using automated matching techniques?}

  All the classification algorithms trialled were able to successfully classify
  web browser traffic routed over the simulated Tor network, with the best
  performing algorithm able to correctly classify 95\%.

  The value of this level of accuracy depends on the application. Identifying
  multiple regular connections to the Tor network may provide sufficient
  confidence to classify a system as a Tor node. This could be combined with
  other known properties of the Tor network to increase confidence.

  \item{Does Tor traffic have characteristics that make it readily distinguishable using
    heuristics based matching?}

  Insert answer here.

  \item{Does Tor traffic have characteristics that make it distinguishable using
    machine learning techniques?}

  Insert answer here.

\end{enumerate*}

\section{Possible Implications of Results}

The research results suggest that the strategies used by anonymous networks to
hide the identity of network participants, may make those anonymous networks
more easy to distinguish from networks and software that do not provide any
anonymity.

Research has shown that the low latency design of Tor means it is vulnerable to
traffic analysis attacks that correlate traffic patterns with users. This
vulnerability is indicative of the subtle amount of overhead incurred when
sending traffic through Tor. This should make it difficult to identify whether a
connection is using Tor VS not using Tor, though it is clearly still possible.
Reducing this overhead further may reduce the success of Tor traffic
identification.

Stronger anonymity providing networks utilize more sophisticated techniques to
hide individual identities such as batching, queuing and normalization. While
they may be more resistant to identity uncovering attacks, this may also make
them more easily identifiable by classification techniques.

\chapter{Conclusion}

\section{Conclusion}

Using Tor as a communications proxy incurs some overhead, sufficient enough that
when using a Tor proxy in controlled conditions, Tor traffic can be
distinguished from regularly encrypted traffic. This overhead may be sufficient
enough that Tor nodes in a real world network can be identified by using readily
available eavesdropping techniques.

However, this experiment was based on a small set of simulated data, with which
it would be impossible to cover all the possible real world conditions. The
variability present in the real Tor network may make this classification
technique impossible. Further research needs to be conducted with live packet
traces from real participants in the Tor network.